Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries

### 2024-04-18 paper.qmd
added some models, neeed to do more cleaning

---
title: "Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries"
author: "Vincent Xie"
toc: true
number-sections: false
execute:
    echo: false
highlight-style: pygments
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: 
    fontsize: 12pt
    geometry: 
      - top=30mm
      - left=20mm
header-includes:
  - \usepackage{setspace}
  - \doublespacing
##  docx: default
---
```{python}
# Packages
import pandas as pd
```

@@

>>>>

---
title: "Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries"
author: "Vincent Xie"
toc: false
number-sections: false
execute:
    echo: false
highlight-style: pygments
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: 
    fontsize: 12pt
    geometry: 
      - top=30mm
      - left=20mm
header-includes:
  - \usepackage{setspace}
  - \doublespacing
##  docx: default
---
```{python}
# Packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
```

@@

# Modeling

```{python}
# Bar chart for frequency of each injury type in InjuryRecord
injury_freq = injury['BodyPart'].value_counts()
plt.figure(figsize=(5, 2))
injury_freq.plot(kind='bar', color='skyblue')
plt.title('Frequency of Each Injury')
plt.xlabel('Body Part')
plt.ylabel('Frequency')
plt.show()

# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(8, 4))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()

# Merge injury_record and playlist on PlayerKey
merged_data = pd.merge(injury, plays, on='PlayKey', how='inner')

# Calculate the frequency of injuries by roster position
injury_position_frequency = merged_data['RosterPosition'].value_counts()

# Plotting the frequency of injuries by roster position
plt.figure(figsize=(8, 4))  # Smaller plot size for better readability
injury_position_frequency.plot(kind='bar', color='salmon')
plt.title('Frequency of Injuries by Roster Position')
plt.xlabel('Roster Position')
plt.ylabel('Frequency')
plt.tight_layout()  # Adjust layout to fit elements
plt.show()
```

```{python}
# Merge injury_record and playlist on PlayerKey
merged_data = pd.merge(injury, plays, on='PlayKey', how='inner')

# Group by 'RosterPosition' and 'BodyPart' and count the occurrences
injury_by_position = merged_data.groupby(['RosterPosition', 'BodyPart']).size().reset_index(name='Frequency')

# Create a pivot table with 'RosterPosition' as index, 'BodyPart' as columns, and 'Frequency' as values
pivot_table = injury_by_position.pivot_table(index='RosterPosition', columns='BodyPart', values='Frequency', fill_value=0)

# Convert frequencies to integers
pivot_table = pivot_table.astype(int)

# Plot a heatmap to visualize the frequency of each type of injury by each position group
plt.figure(figsize=(14, 8))
sns.heatmap(pivot_table, annot=True, cmap="viridis")
plt.title('Frequency of Each Type of Injury by Position Group')
plt.xlabel('Body Part')
plt.ylabel('Roster Position')
plt.show()
```
